---
title: "Tak 4 - Halloween Candy"
output: html_notebook
---

 
  # ============================================================================
  
  #  ----- File Name: halloween_candy_analysis.Rmd
  #  ----- Description: Analysis of BoingBoing Halowwen Candy Data
  #  ----- Output: 
  #  ----- Created Date: 30th Nov 2022
  #  ----- Created By: Sarah Hughes
  
  # ============================================================================
  
  #  ----- Assumptions: 
   
  # ============================================================================
  
  #  ----- Version History: v0.1
  
  # ============================================================================
  
  #  ----- Peer reviewer:
  
  #  ----- Peer review date:
  
  # ============================================================================
  

```{r Load libraries}
# load libraries
library(tidyverse)
library(skimr)
library(assertr)
library(here)
library(testthat)
```
`
```{r Data Cleaning Script}
# call data cleaning script
source(here("data_cleaning_scripts","clean_candy_data.R")) 
```
```{r Load Cleaned data}
# read in cleaned data
clean_candy_data <- read_csv(here("clean_data","clean_candy_data.csv"))
```

### Q1 What is the total number of candy ratings given across the three years.
```{r}
clean_candy_data %>% 
  filter(!(is.na(rating))) %>% 
  group_by(rating) %>% 
  summarise(count = n())
```
```{r}
clean_candy_data %>% 
  filter(!(is.na(rating))) %>% 
  summarise(count = n())
```
### Q2 What was the average age of people who are going out trick or treating?
### Q3 What was the average age of people who are NOT going trick or treating?
```{r}
clean_candy_data %>% 
  select(tot_demog_q) %>% 
  group_by(tot_demog_q) %>% 
  summarise(count = n())
```
```{r}
clean_candy_data %>% 
  select(tot_demog_q, age_clean) %>% 
  group_by(tot_demog_q) %>% 
  summarise(avg_age = mean(age_clean, na.rm=TRUE))
```
### Q4 For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
clean_candy_data %>% 
  select(candy_type,rating) %>% 
  filter(rating == "MEH") %>% 
  group_by(candy_type,rating) %>% 
  summarise(num_ratings = n()) %>% 
  arrange(desc(num_ratings)) %>% 
  head(1)
```

```{r}
clean_candy_data %>% 
  select(candy_type,rating) %>% 
  filter(rating == "JOY") %>% 
  group_by(candy_type,rating) %>% 
  summarise(num_ratings = n()) %>% 
  arrange(desc(num_ratings)) %>% 
  head(1)
```
```{r}
clean_candy_data %>% 
  select(candy_type,rating) %>% 
  filter(rating == "DESPAIR") %>% 
  group_by(candy_type,rating) %>% 
  summarise(num_ratings = n()) %>% 
  arrange(desc(num_ratings)) %>% 
  head(1)
```
### Q5 How many people rated Starburst as despair?
```{r}
clean_candy_data %>% 
  select(candy_type,rating) %>% 
  filter(candy_type == "starburst") %>% 
  group_by(rating) %>% 
  summarise(num_ratings = n()) 
```

### create a numeric column for rating
```{r}
clean_candy_data <- clean_candy_data %>% 
  mutate(rating_num = case_when(
    rating == "DESPAIR" ~ as.numeric(-1),
    rating == "MEH" ~ as.numeric(0),
    rating == "JOY" ~ as.numeric(1),
    TRUE ~ NA_real_))
```

### Q6 Most popular candy rating
```{r}
clean_candy_data %>% 
  select(gender_demog_q,candy_type,rating_num) %>% 
  group_by(gender_demog_q,candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(gender_demog_q,candy_type) %>% 
  head(10)
```
### Q7 Most popular candy rating
```{r}
clean_candy_data %>% 
  select(year,candy_type,rating_num) %>% 
  group_by(year) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(10)
```
### Q8 Most popular candy rating
```{r}
clean_candy_data %>% 
  filter(!(is.na(country_clean))) %>% 
  select(country_clean,candy_type,rating_num) %>% 
  group_by(country_clean,candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(10)
```
```{r}
clean_candy_data %>% 
  select(candy_type) %>%
  arrange(desc(candy_type)) %>% 
  distinct(candy_type)
```
