---
title: "Task 4 - Halloween Candy"
output: html_notebook
---

=================================================================

\# ----- File Name: halloween_candy_analysis.Rmd \# ----- Description: Analysis
of boingboing halloween candy data. This data is collected in an annual survey,
the purpose of the survey is to determine which candy is most popular by asking
the question. This analysis combines data from the years 2015, 2016 and 2017.
The dataset contains demographic data for each rater: age, gender, country and
'going out trick or treating' intention. Demographic data was an optional field
and so there are some missing data. Note that country and gender data were not
collected in 2015.

Candy has been rated using a 3 point scale from 'despair-meh-joy'. A numeric
rating for each candy has been calculated using the formulas despair = 1, meh =0
and joy =-1. 

\# ----- Output: \# ----- Created Date: 30th Nov 2022 \# ----- Created By: Sarah
Hughes

=================================================================

\# ----- Assumptions:

Age Data

Rater ages have been marked as NA if they fell outside expected limits, 
minimum age =1 and maximum age = 120. If an age range was given than an assumed age was calculated as the average of the range. If an approximate age was given (e.g. 40's, 50+ or 30ish) the age was assumed to be the decade + 5. Where age was recorded as 'old', age was set to 75.

Country Data

In the analysis, all countries not identified as either usa, canada, uk or usa
have been assigned to 'other'. This includes some entries that don't match known
countries.

Candy Data

The following candy types have been identified as equivalent

1.  boxo_raisins = box_o\_raisins

2.  dark_chocolate_hershey = hersheys_dark_chocolate

3.  100_grand_bar = x100_grand_bar

4.  bonkers = bonkers_the_candy

5.  sweetums_a\_friend_to_diabetes = sweetums

6.  hershey_s\_kissables = hersheys_kisses

7.  licorice = licorice_yes_black

=====================================================

\# ----- Version History: v0.1

=====================================================

\# ----- Peer reviewer:

\# ----- Peer review date:

=====================================================

```{r Load libraries}
# load libraries
library(tidyverse)
library(here)
```

```{r Data Cleaning Script}
# if not already run call data cleaning script here
# source(here("data_cleaning_scripts","clean_candy_data.R")) 
```

```{r Load Cleaned data}
# read in cleaned data
clean_candy_data <- read_csv(here("clean_data","clean_candy_data.csv"))
clean_rater_data <- read_csv(here("clean_data","clean_rater_data.csv"))

clean_candy_all<-  left_join(clean_candy_data,clean_rater_data)
```

```{r Summary}
clean_candy_all %>% 
  select(rater_id) %>% 
  distinct(rater_id)

clean_candy_all %>% 
  select(candy_type) %>% 
  distinct(candy_type)
```





```{r}
clean_candy_all %>% 
  group_by(rating) %>% 
  summarise(count = n())
```

```{r}
all_ratings <- clean_candy_all %>% 
  summarise(count = n())
```

T
```{r}
clean_rater_data %>% 
  select(tot_clean, age_clean) %>% 
  group_by(tot_clean) %>% 
  summarise(avg_age = mean(age_clean, na.rm=TRUE), count=n())
```

```{r}
clean_candy_all %>% 
  select(candy_type,rating) %>% 
  filter(rating == "joy") %>% 
  group_by(candy_type,rating) %>% 
  summarise(num_ratings = n(), 
            percent_rated = 100*(num_ratings/nrow(clean_rater_data))) %>% 
  arrange(desc(num_ratings)) %>% 
  head(3)
```

The candy bar 'any_full_sized_candy_bar' received the most 'joy' ratings. A
total of 7589 ratings, from 88% of raters.

```{r}
clean_candy_all %>% 
  select(candy_type,rating) %>% 
  filter(rating == "meh") %>% 
  group_by(candy_type,rating) %>% 
  summarise(num_ratings = n(), 
            percent_rated = 100*(num_ratings/nrow(clean_rater_data))) %>% 
  arrange(desc(num_ratings)) %>% 
  head(3)
```

'lollipops' received the most 'meh' ratings, with 1570 in total. 18% of raters
gave this rating over the three years

```{r}
clean_candy_all %>% 
  select(candy_type, rating) %>% 
  filter(rating == "despair") %>% 
  group_by(candy_type, rating) %>% 
  summarise(num_ratings = n(), 
            percent_rated = 100*(num_ratings/nrow(clean_rater_data))) %>% 
  arrange(desc(num_ratings)) %>% 
  head(3)
```



```{r}
clean_candy_all %>% 
  select(candy_type, rating) %>% 
  filter(candy_type == "starburst", rating == "despair") %>% 
  group_by(rating) %>% 
  summarise(num_ratings = n()) 
```


```{r}
clean_candy_all <- clean_candy_all %>% 
  mutate(rating_num = case_when(
    rating == "despair" ~ as.numeric(-1),
    rating == "meh" ~ as.numeric(0),
    rating == "joy" ~ as.numeric(1),
    TRUE ~ NA_real_))
```


```{r}
clean_rater_data %>% 
   select(gender_clean) %>%
   group_by(gender_clean) %>% 
   summarise(count = n())
```

```{r}
clean_candy_all %>% 
  select(gender_clean, candy_type, rating_num) %>% 
  filter(gender_clean == "male") %>% 
  group_by(gender_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(3)
```

```{r}
clean_candy_all %>% 
  select(gender_clean, candy_type, rating_num) %>% 
  filter(gender_clean == "female") %>% 
  group_by(gender_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(3)
```

```{r}
clean_candy_all %>% 
  select(gender_clean, candy_type, rating_num) %>% 
  filter(gender_clean == "prefer not to say") %>% 
  group_by(gender_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(3)
```

```{r}
clean_candy_all %>% 
  select(gender_clean, candy_type, rating_num) %>% 
  filter(gender_clean == "other") %>% 
  group_by(gender_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(3)
```
```{r}
clean_candy_all %>% 
  select(gender_clean, candy_type, rating_num) %>% 
  filter(is.na(gender_clean)) %>% 
  group_by(gender_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(3)
```

```{r}
clean_rater_data %>% 
   group_by(year) %>% 
   summarise(count = n())
```

```{r}
clean_candy_all %>% 
  select(year,candy_type,rating_num) %>% 
  group_by(candy_type,year) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  filter(year == 2015) %>% 
  arrange(desc(candy_ratings),year) %>% 
  head(1)
```

```{r}
clean_candy_all %>% 
  select(year,candy_type,rating_num) %>% 
  group_by(candy_type,year) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  filter(year == 2016) %>% 
  arrange(desc(candy_ratings),year) %>% 
  head(1)
```

```{r}
clean_candy_all %>% 
  select(year,candy_type,rating_num) %>% 
  group_by(candy_type,year) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  filter(year == 2017) %>% 
  arrange(desc(candy_ratings),year) %>% 
  head(1)
```

```{r}
#check 2015 data
clean_candy_all %>% 
  filter(year==2015) %>% 
  group_by(rating) %>% 
  summarise(count = n())
```

```{r}
clean_rater_data %>% 
   group_by(country_clean) %>% 
   summarise(count = n())
```

```{r}
clean_candy_all %>% 
  filter(country_clean=="usa") %>% 
  select(country_clean,candy_type,rating_num) %>% 
  group_by(candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(1)
```

```{r}
clean_candy_all %>% 
  filter(country_clean=="canada") %>% 
  select(country_clean,candy_type,rating_num) %>% 
  group_by(candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(desc(candy_ratings)) %>% 
  head(1)
```

```{r}
clean_candy_all %>% 
  filter(country_clean=="uk") %>% 
  select(country_clean,candy_type,rating_num) %>% 
  group_by(country_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(country_clean,desc(candy_ratings)) %>% 
  head(10)
```
```{r}
clean_candy_all %>% 
  filter(country_clean=="other") %>% 
  select(country_clean,candy_type,rating_num) %>% 
  group_by(country_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(country_clean,desc(candy_ratings)) %>% 
  head(10)
```
```{r}
clean_candy_all %>% 
  filter(is.na(country_clean)) %>% 
  select(country_clean,candy_type,rating_num) %>% 
  group_by(country_clean, candy_type) %>% 
  summarise(candy_ratings = sum(rating_num, na.rm= TRUE)) %>% 
  arrange(country_clean,desc(candy_ratings)) %>% 
  head(10)
```
```{r}
clean_candy_all %>% 
  select(candy_type) %>%
  arrange(desc(candy_type)) %>% 
  distinct(candy_type)
```
