---
title: "Workbook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
library(here)
library(skimr)
library(janitor)
```


```{r}
ship_data  <- read_excel(here::here("raw_data/seabirds.xls"), sheet = "Ship data by record ID")
bird_data  <- read_excel(here::here("raw_data/seabirds.xls"), 
                         sheet = "Bird data by record ID",
                         col_type = c("text"))
# error reading sex column as logical

#tidy up variable names
bird_clean <- clean_names(bird_data) 
ship_clean <- clean_names(ship_data) 
```

```{r}
names(bird_clean)
names(ship_clean)
```

```{r}
# check occurrence data - there can be a mismatch between the count and the specific numbers
# not helpful for filling in gaps in count
bird_clean %>% 
  select(count, nfeed, nsow, nsoice, nflyp, nacc, nfoll) %>% 
  mutate(across(c(nfeed,nsow,nsoice,nflyp,nacc,nfoll), as.numeric)) %>% 
  mutate(count_check = rowSums(across(c(nfeed,nsow,nsoice,nflyp, nacc, nfoll)))) %>% 
  select(count, count_check, nfeed, nsow, nsoice, nflyp, nacc, nfoll) %>% 
  filter((is.na(count) & (count_check>0)))
```


** Notes**
In this analysis we are interested 
in bird names as scientific name, common name and species abbrev. 
also sighting info time, date, lat and long?

First remove columns we don't need and rename columns for ease of use

```{r}
ship_clean <- ship_clean %>% 
  select(record, 
         record_id,
         date,
         time,
         lat,
         long, 
         long360)
```


```{r}
bird_clean <- bird_clean %>% 
  select(record, 
         record_id,
         species_common_name_taxon_age_sex_plumage_phase, 
         species_scientific_name_taxon_age_sex_plumage_phase,
         species_abbreviation, 
         # following columns are needed for cleaning
         age,
         wanplum,
         plphase,
         sex,
         count) %>% 
  
  # rename variables to make them easier to work with
  rename(species_common_name = 
           species_common_name_taxon_age_sex_plumage_phase) %>% 
  rename(species_science_name =
           species_scientific_name_taxon_age_sex_plumage_phase)
```

```{r}
skim(bird_clean)
```
```{r}
skim(ship_clean)
```
```{r}
bird_clean %>% 
  distinct(species_abbreviation)
```

```{r}
# recode [NO BIRDS RECORDED] as NA 
bird_clean <- bird_clean %>%
   mutate(species_common_name = na_if(species_common_name,"[NO BIRDS RECORDED]")) %>% 
   drop_na(species_common_name)
```

```{r}
# original thought to code PL6 as an error
#later decided to allow it as its being removed anyway - assume code in wanplum column is correct
#bird_clean %>% 
#   mutate(species_common_clean = if_else(
#     str_sub(species_common_name, -3, -1) == "PL6",
#     NA, species_common_name)
#       ))
```
```{r}
# fix coding errors - also no longer needed as changed strategy
# this will be removed later anyway but its the right way to remove it!
#bird_clean %>% 
#   mutate(species_common_clean =    #str_replace(species_common_name,"PL6","PL5")
#       ) %>%
#   mutate(species_science_clean =    #str_replace(species_science_name,"PL6","PL5")
#       ) %>%
#   mutate(species_abbrev_clean=    #str_replace(species_abbreviation,"PL6","PL5")
#       ) %>%  
#filter(str_detect(species_common_name,"PL6")) %>% 
#select(species_common_name, species_common_clean, wanplum)         
```



```{r}
#take a look at species info
bird_clean %>% 
  distinct(species_common_name) %>% 
  arrange(species_common_name)
```
```{r}
# set numeric variables back
bird_clean <- bird_clean %>% 
  mutate(record_id = as.numeric(record_id)) %>% 
  mutate(record = as.numeric(record)) %>% 
  mutate(count = as.numeric(count))
```
```{r}
#take a look at species info
bird_clean %>%
  select(age, species_abbreviation) %>% 
  filter(!is.na(age)) %>% 
  arrange(species_abbreviation)
```
# species name contains additional info on age, sex, plumage phase that is not needed for this analyis
# assume we are just interested in the species
# this additional data can be removed into a seperate column

```{r}
# test how we extract info
#names(bird_clean)
```


** Notes
looks like ages, sex and plumage columns can hep identify additional data to be removed from species names
?? sensu lato

```{r}
# original version
#age_codes <- distinct(bird_clean, age) %>% 
#drop_na(age) %>% 
#  pull()
#plum_codes <- distinct(bird_clean, wanplum) %>% 
#drop_na(wanplum) %>% 
#pull()
#plum_codes <- str_c("PL",plum_codes)
#phase_codes <- distinct(bird_clean, plphase) %>% 
#drop_na(plphase) %>% 
#pull()  
#sex_codes <- distinct(bird_clean, sex) %>% 
#drop_na(sex) %>% 
#pull()
```
```{r}
#an easier way is to identify any additional codes from species_abbreviation
extra_codes <- bird_clean %>%
  select(species_abbreviation) %>% 
  separate(species_abbreviation, 
           c("spec_abbrev", "extra1","extra2")," ")
# union finds unique values 
combined_codes <- as_tibble (union_all(extra_codes$extra1,extra_codes$extra2)) %>% 
  drop_na() %>% 
  distinct() %>% 
  pull()
# some birds are coded with more than one identifier
# extracting extra 1 and extra 2 allows for case where unique info could be captured this way
# as it is  - nothing new comes from extra2 its just LGHT and DRK
# this code would fail if there were three identifiers. Is it possible to count spaces and so identify how many columns needed. Mayb leavee for another task!

# this captures the problematic coding of ADF and PL6 which were left behind by previous attempt
  

```
```{r}
#bird_clean %>% 
#distinct(age)
#distinct(spec_abbrev_add1)
```

```{r}
# keeps going until all strings have been removed
strip_strings <- function(old_column, strings){
  cleaned <- old_column
  
  # order strings longest first so "string within string" is not a problem
  order <- sort(nchar(strings), decreasing=TRUE, index.return=TRUE)
  strings <- strings[order$ix]

  for (badstring in strings){
    cleaned <- str_remove(cleaned,str_c(" ",badstring)) #adding this doesn't allow for a string to be 'embedded'
    #in this case it fixes problem of accidentally removing 'F' and 'M' at the start of species name
    cleaned <- str_trim(cleaned)
  }
    return(cleaned)  
}
```


```{r}
#testing use of string_remove_all
testdata <-  c("dog AA","catCC BB","BBhorse AAB","big mouse AA BB") 
testdata1 <- str_remove_all(testdata,c("AA","BB","CC","AAB"))
testdata1
#doesn't capture all instances of all strings - maybe because I am not using it correctly?


#test strip_strings function
cleandata <-  strip_strings(testdata,c("AA","BB","CC","AAB"))
cleandata
```

```{r}
#now use codes to clean species_common_name
#bird_clean <- bird_clean %>% 
#  mutate(species_common_clean = strip_strings(species_common_name, #age_codes)) %>% 
#  mutate(species_common_clean = strip_strings(species_common_clean, #plum_codes)) %>% 
#  mutate(species_common_clean = strip_strings(species_common_clean, #phase_codes)) %>% 
#  mutate(species_common_clean = strip_strings(species_common_clean, #sex_codes)) 
```
```{r}
#now use codes to clean species_scientific_name
#bird_clean <- bird_clean %>% 
#  mutate(species_science_clean = strip_strings(species_science_name, #age_codes)) %>% 
#  mutate(species_science_clean = strip_strings(species_science_clean, #plum_codes)) %>% 
#  mutate(species_science_clean = strip_strings(species_science_clean, #phase_codes)) %>% 
#  mutate(species_science_clean = strip_strings(species_science_clean, #sex_codes)) 
```

```{r}
#now use codes to clean species_abbreviation
#bird_clean <- bird_clean %>% 
#  mutate(species_abbrev_clean = strip_strings(species_abbreviation, #age_codes)) %>% 
#  mutate(species_abbrev_clean = strip_strings(species_abbrev_clean, #plum_codes)) %>% 
#  mutate(species_abbrev_clean = strip_strings(species_abbrev_clean, #phase_codes)) %>% 
#  mutate(species_abbrev_clean = strip_strings(species_abbrev_clean, #sex_codes)) 
```

```{r}
#now use codes to clean species_common_name
bird_clean <- bird_clean %>% 
  mutate(species_common_clean = strip_strings(species_common_name, combined_codes)) %>% 
  mutate(species_science_clean = strip_strings(species_science_name, combined_codes)) %>% 
  mutate(species_abbrev_clean = strip_strings(species_abbreviation, combined_codes)) 
```

```{r}
# 1 case of ADF
bird_clean %>% 
  select(species_science_name,age,sex,species_science_clean) %>% 
  filter(str_detect(species_science_name, "ADF"))

```

```{r}
# 1 case of PL6
bird_clean %>% 
  select(species_science_name,wanplum,species_science_clean) %>% 
  filter(str_detect(species_science_name, "PL6"))
```


```{r}
bird_clean %>% 
  select(age, species_common_clean) %>% 
  filter(!is.na(age)) %>% 
  group_by(age) %>% 
  mutate(count = n())
```
```{r}
bird_clean %>% 
  select(species_common_clean, species_science_clean, species_abbrev_clean) %>% 
  filter(str_detect(species_abbrev_clean, "PEN")) 
  #distinct(species_common_clean,species_science_clean,species_abbrev_clean) 
```
```{r}
bird_clean %>% 
  select(species_common_clean, species_science_clean, species_abbrev_clean,count) %>% 
  filter(str_detect(species_common_clean, "penguin")) 
  #distinct(species_common_clean,species_science_clean,species_abbrev_clean) 
```

```{r}
bird_clean %>% 
  select(species_common_clean, species_science_clean, species_abbrev_clean) %>% 
  filter(is.na(species_science_clean)) %>% 
  distinct(species_common_clean,species_science_clean,species_abbrev_clean) 
```

```{r}
bird_clean %>% 
  select(species_common_clean, species_science_clean, species_abbrev_clean) %>% 
  filter(str_detect(species_common_clean, "penguin")) %>% 
  distinct(species_common_clean,species_science_clean,species_abbrev_clean) 
```


```{r}
bird_clean %>% 
  select(species_common_clean, species_science_clean, species_abbrev_clean) %>% 
  group_by(species_abbrev_clean,species_common_clean) %>% 
  mutate(count =  n())
```
** notes **
now need to look at the different ways general species have been coded.
search on sp or sp. in science name and do some recoding.
sensu lato -- equivalent to (unidentified)
sp should be corrected to sp.


```{r}
skim(bird_clean)
```
```{r}
bird_clean %>% 
  select(species_common_clean,species_science_clean,species_abbrev_clean) %>% 
    mutate(count = coalesce(count, 1) 
           str_replace(species_common_clean,"sensu lato","(unidentified)")) 
```

```{r}
bird_clean %>% 
  select(species_common_clean,species_science_clean,species_abbrev_clean) %>% 
  filter(str_detect(species_science_clean, "sp")) %>% 
    mutate(species_science_clean = 
           str_replace(species_science_clean,"sp.","sp")) %>% 
    mutate(species_science_clean = 
           str_replace(species_science_clean,"sp","sp."))  
```
```{r}
bird_clean <- bird_clean %>% 
    mutate(species_common_clean = 
           str_replace(species_common_clean,"sensu lato","(unidentified)")) %>% 
    mutate(count = coalesce(count, 1))
```
```{r}
#join datasets
all_bird_data <- left_join(bird_clean, ship_clean, 'record_id') %>% 
```



# highest total count

```{r}
all_bird_data %>% 
  select(species_common_clean, species_science_clean, species_abbrev_clean, count, lat) %>% 
  group_by(species_abbrev_clean) %>% 
  mutate(total_num = sum(count)) 
```





```{r}
ship_clean %>% 
  distinct(ew)
```

```{r}
seabirds_clean <- seabirds_raw %>%
  select(record_id,lat,long360, )
  mutate(long2 = if_else(ew == "W", -1*long,long))
seabirds_clean
```
 