---
title: "Workbook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
library(here)
library(skimr)
library(janitor)
```


```{r}
ship_data  <- read_excel(here::here("raw_data/seabirds.xls"), sheet = "Ship data by record ID")
bird_data  <- read_excel(here::here("raw_data/seabirds.xls"), 
                         sheet = "Bird data by record ID",
                         col_type = c("text"))
# error reading sex column as logical

#tidy up variable names
bird_clean <- clean_names(bird_data) 
ship_clean <- clean_names(ship_data) 
```

```{r}
names(bird_clean)
names(ship_clean)
```

** Notes**
In this analysis we are interested 
in bird names as scientific name, common name and species abbrev. 
also sighting info time, date, lat and long?

First remove columns we don't need and rename columns for ease of use

```{r}
ship_clean <- ship_clean %>% 
  select(record, 
         record_id,
         date,
         time,
         lat,
         long, 
         long360)
```


```{r}
bird_clean <- bird_clean %>% 
  select(record, 
         record_id,
         species_common_name_taxon_age_sex_plumage_phase, 
         species_scientific_name_taxon_age_sex_plumage_phase,
         species_abbreviation, 
         # following columns are needed for cleaning
         age,
         wanplum,
         plphase,
         sex,
         count) %>% 
  
  # rename variables to make them easier to work with
  rename(species_common_name = 
           species_common_name_taxon_age_sex_plumage_phase) %>% 
  rename(species_science_name =
           species_scientific_name_taxon_age_sex_plumage_phase)
```

```{r}
skim(bird_clean)
```
```{r}
skim(ship_clean)
```
```{r}
bird_clean %>% 
  distinct(sex)
```

```{r}
# recode [NO BIRDS RECORDED] as NA 
bird_clean <- bird_clean %>%
   mutate(species_common_name = na_if(species_common_name,"[NO BIRDS RECORDED]")) %>% 
   drop_na(species_common_name)
```



```{r}
#take a look at species info
bird_clean %>% 
  distinct(species_common_name) %>% 
  arrange(species_common_name)
```
```{r}
# set numeric variables back
bird_clean <- bird_clean %>% 
  mutate(record_id =  as_numeric(record_id)) %>% 
  mutate(record = as.numeric(record))       
```
```{r}
#take a look at species info
bird_clean %>%
  select(age, species_abbreviation) %>% 
  filter(!is.na(age)) %>% 
  arrange(species_abbreviation)
```
# species name contains additional info on age, sex, plumage phase that is not needed for this analyis
# assume we are just interested in the species
# this additional data can be removed into a seperate column

```{r}
# test how we extract info
#names(bird_clean)

#bird_clean <- bird_clean %>% 
#  separate(species_abbreviation, c("spec_abbrev", #"spec_abbrev_add1","spec_abbrev_add2")," ")
#```
```{r}
#bird_clean %>% 
#distinct(age)
#distinct(spec_abbrev_add1)
```

** Notes
looks like ages, sex and plumage columns can hep identify additional data to be removed from species names
?? sensu lato

```{r}
age_codes <- distinct(bird_clean, age) %>% 
drop_na(age) %>% 
  pull()
plum_codes <- distinct(bird_clean, wanplum) %>% 
drop_na(wanplum) %>% 
pull()
plum_codes <- str_c("PL",plum_codes)
phase_codes <- distinct(bird_clean, plphase) %>% 
drop_na(plphase) %>% 
pull()  
sex_codes <- distinct(bird_clean, sex) %>% 
drop_na(sex) %>% 
pull()
```

```{r}
# keeps going until all strings have been removed
strip_strings <- function(old_column, strings){
  cleaned <- old_column
  
  # order strings longest first so "string within string" is not a problem
  order <- sort(nchar(strings), decreasing=TRUE, index.return=TRUE)
  strings <- strings[order$ix]

  for (badstring in strings){
    cleaned <- str_remove(cleaned,badstring)
    cleaned <- str_trim(cleaned)
  }
    return(cleaned)  
}
```


```{r}
#testing use of string_remove_all
testdata <-  c("dog AA","catCC BB","hBBorse AAB","big mouse AA BB") 
testdata1 <- str_remove_all(testdata,c("AA","BB","CC","AAB"))
testdata1
#doesn't capture all instances of all strings - maybe because I am not using it correctly?

#test strip_strings function
cleandata <-  strip_strings(testdata,c("AA","BB","CC","AAB"))
cleandata
```

```{r}
#now use codes to clean species_common_name
bird_clean <- bird_clean %>% 
  mutate(species_common_clean = strip_strings(species_common_name, age_codes)) %>% 
  mutate(species_common_clean = strip_strings(species_common_clean, plum_codes)) %>% 
  mutate(species_common_clean = strip_strings(species_common_clean, phase_codes)) %>% 
  mutate(species_common_clean = strip_strings(species_common_clean, sex_codes)) 
```
```{r}
#now use codes to clean species_scientific_name
bird_clean <- bird_clean %>% 
  mutate(species_science_clean = strip_strings(species_science_name, age_codes)) %>% 
  mutate(species_science_clean = strip_strings(species_science_clean, plum_codes)) %>% 
  mutate(species_science_clean = strip_strings(species_science_clean, phase_codes)) %>% 
  mutate(species_science_clean = strip_strings(species_science_clean, sex_codes)) 
```

```{r}
#now use codes to clean species_scientific_name
bird_clean <- bird_clean %>% 
  mutate(species_abbrev_clean = strip_strings(species_abbreviation, age_codes)) %>% 
  mutate(species_abbrev_clean = strip_strings(species_abbrev_clean, plum_codes)) %>% 
  mutate(species_abbrev_clean = strip_strings(species_abbrev_clean, phase_codes)) %>% 
  mutate(species_abbrev_clean = strip_strings(species_abbrev_clean, sex_codes)) 
```



```{r}
bird_clean %>% 
  select(age, species_common_clean) %>% 
  filter(!is.na(age)) %>% 
  group_by(age) %>% 
  mutate(count = n())
```



```{r}
bird_clean %>% 
distinct(species_common_clean) %>% 
arrange(species_common_clean)
```



```{r}
ship_clean %>% 
  distinct(ew)
```

```{r}
seabirds_clean <- seabirds_raw %>%
  select(record_id,lat,long360, )
  mutate(long2 = if_else(ew == "W", -1*long,long))
seabirds_clean
```
 
