---
title: "Task 3 - Seabird"
output: html_notebook
---
  
  # ============================================================================
  
  #  ----- File Name: seabird_analysis.Rmd
  #  ----- Description: Analysis of seabird data
  #  ----- Output: Analysis of seabird data, examining rate of observation
     and abundance of various seabirds
  #  ----- Created Date: 28th Nov 2022
  #  ----- Created By: Sarah Hughes
  
  # ============================================================================
  
  #  ----- Assumptions: 
    Where there is species data in a record but the species count is 'NA' 
    sets abundance as 1, e.g at least 1 bird was observed.
  
  # ============================================================================
  
  #  ----- Version History: v0.1
  
  # ============================================================================
  
  #  ----- Peer reviewer:
  
  #  ----- Peer review date:
  
  # ============================================================================
  

```{r Load libraries}
# load libraries
library(tidyverse)
library(skimr)
library(assertr)
library(here)
library(testthat)
```
`
```{r Data Cleaning Script}
# call data cleaning script
source(here("data_cleaning_scripts","cleaning_bird_data.R")) 
```

```{r Load Cleaned data}
# read in cleaned data
clean_data <- read_csv(here("clean_data","joined_bird_data.csv"))
```
### Q1 Which bird has most individual sightings? 
### One way to interpret this is which bird was seen in the biggest group on one individual sighting?

```{r}
joined_bird_data %>% 
  select(species_common_clean,species_science_clean,species_abbrev_clean,abundance) %>% 
  arrange(desc(abundance)) %>% 
  distinct(species_common_clean,species_science_clean,species_abbrev_clean,abundance) %>% 
  #recode high abundance for clarity in summary table
  mutate(abundance = as.character(abundance)) %>% 
  mutate(abundance = recode(abundance, "99999" = ">100,000")) %>% 
  head(5)
```
### Both the Short-tailed shearwater and the Prion were spotted in groups of greater than 100,000

### Q1 Which bird had the most individual sightings? 
### Another way is to count number of individual records for each type of bird and find the highest
```{r} 
joined_bird_data %>% 
  group_by(species_common_clean) %>% 
  mutate(sightings_count = n()) %>% 
  arrange(desc(sightings_count)) %>% 
  distinct(species_common_clean,species_science_clean,species_abbrev_clean,sightings_count) %>% 
  head(1)
```
### The wandering albatross was observed in 11293 individual sighting records
```{r}
#check this
joined_bird_data %>% 
  filter(species_abbrev_clean == "DIOANTEXU") %>% 
  nrow
```

### Q2 Which bird had the highest total abundance
```{r}
joined_bird_data %>% 
  group_by(species_common_clean,species_science_clean,species_abbrev_clean) %>% 
  summarise(total_bird_abundance = sum(abundance)) %>%
  arrange(desc(total_bird_abundance)) %>% 
  head(5)
```
### The bird seen with highest abundance was the Short-tailed shearwater, more than 982,669 were seen in total
    

### Q3 Which bird had the highest total count above latitude -30
```{r}
joined_bird_data %>% 
  filter(lat >-30) %>% 
  group_by(species_common_clean,species_science_clean,species_abbrev_clean) %>% 
  summarise(total_bird_abundance = sum(abundance)) %>%
  arrange(desc(total_bird_abundance)) %>% 
  head(5)
```
### The Wedge-tailed shearwater was observed in 855 individual sighting records

### Q4 How many birds were seen only in groups of one?
```{r}
joined_bird_data %>% 
  filter(abundance == 1) %>% 
  group_by(species_common_clean,species_science_clean,species_abbrev_clean) %>%
  distinct(species_common_clean) %>% 
  nrow
```
### 150 different types of birds were only seen in groups of 1 

### Q5 First checl how many types of penguins were seen?
```{r}

joined_bird_data %>% 
  filter(str_detect(str_to_lower(species_common_clean), "penguin")) %>% 
  group_by(species_common_clean,species_science_clean,species_abbrev_clean) %>% 
  summarise(count_penguin_obs =  n(), abundance_penguins = sum(abundance)) %>% 
  arrange(desc(count_penguin_obs))
```
### There were 7 different types of penguins

### Q5 How many total penguins were seen?
```{r}
joined_bird_data %>% 
   filter(str_detect(str_to_lower(species_common_clean), "penguin"))  %>% 
    summarise(total_abundance_penguins = sum(abundance)) 
```
### 162 penguins seen in total